# RRBS analysis workflow for MethylKit

```{sh modules}
module load R/x86_64/2.15.1
```

```{r setup}
library(methylKit)
library("graphics")
```

Read in the files.
These files are generated by methylation calling using BSSeeker2 which produces a .wig file. These are then converted to bed files suited for MethylKit using a custom script.
```{r input, cache=TRUE}
if(context != ""){
  ext <- paste(".sort.clip.call.",context,".CGmap.", chromosome, ".bed", sep="")
}else{
  ext <- paste(".sort.clip.call.CGmap.", chromosome, ".bed", sep="")
}

file.list <- list(paste("Sample_R06_GCCAAT_L005", ext, sep=""), 
paste("Sample_R07_CAGATC_L003", ext, sep=""), 
paste("Sample_R08_ACTTGA_L005", ext, sep=""), 
paste("Sample_R09_GATCAG_L007", ext, sep=""), 
paste("Sample_R10_TAGCTT_L005", ext, sep=""), 
paste("ICI001A_CGATGT_L001", ext, sep=""), 
paste("ICI001B_TGACCA_L001", ext, sep=""), 
paste("ICI001C_ACAGTG_L001", ext, sep=""), 
paste("ICI002A_ATCACG_L002", ext, sep=""),
paste("ICI002B_TTAGGC_L002", ext, sep=""))

if(context != ""){
  if(context == "CG"){
    rcontext = "CpG"
  }else{
    rcontext=context
  }
  myobj <- read(file.list, sample.id = list("C-146", "C-148", "C-156", "C-157", "C-165","LR5007-172","LR5007-186","LR5007-191","LR5007-171","LR5007-179"), assembly = "rice", treatment = c(1,1,1,1,1,0,0,0,0,0),context=rcontext)
}else{
  myobj <- read(file.list, sample.id = list("C-146", "C-148", "C-156", "C-157", "C-165","LR5007-172","LR5007-186","LR5007-191","LR5007-171","LR5007-179"), assembly = "rice", treatment = c(1,1,1,1,1,0,0,0,0,0))
}
```

Determining methylation stats
```{r methStats}
for (i in 1:length(file.list)){
  getMethylationStats(myobj[[i]], plot = T, both.strands = F)
  getCoverageStats(myobj[[i]], plot = T, both.strands = F)
}
```

Filtering coverage
```{r filter, cache=TRUE}
filtered.myobj <- filterByCoverage(myobj, lo.count = 10, lo.perc = NULL, hi.count = NULL, hi.perc = NULL)
myobj <- filtered.myobj
```

Normalising coverage
```{r normalise, cache=TRUE}
normalised.myobj <- normalizeCoverage(myobj)
myobj <- normalised.myobj
```

Merge all data and determine correlation
```{r merge, cache=TRUE}
meth <- unite(myobj, destrand = FALSE)
head(meth)
if(context==""){
  write.table(meth, file="unite.txt", sep="\t", quote=F)
}else{
  write.table(meth, file=paste("unite.",context,".txt",sep=""), sep="\t", quote=F)
}

getCorrelation(meth, plot = T)
clusterSamples(meth, dist = "correlation", method = "ward", plot = TRUE)
PCASamples(meth, screeplot = TRUE)
PCASamples(meth)
```

Differential methylation
```{r diff, cache=TRUE}
myDiff <- calculateDiffMeth(meth, num.cores = 2)
qvalue <- 0.01
diff <- 25

myDiff25p.hyper <- get.methylDiff(myDiff, difference = diff, qvalue = qvalue, type = "hyper")
myDiff25p.hypo <- get.methylDiff(myDiff, difference = diff, qvalue = qvalue, type = "hypo")
myDiff25p <- get.methylDiff(myDiff, difference = diff, qvalue = qvalue)

if(context==""){
  write.table(myDiff25p.hyper, file="diff.hyper.txt", sep="\t", quote=F)
  write.table(myDiff25p.hypo, file="diff.hypo.txt", sep="\t", quote=F)  
  write.table(myDiff25p, file="diff.all.txt", sep="\t", quote=F)
}else{
  write.table(myDiff25p.hyper, file=paste("diff.hyper.",context,".txt",sep=""), sep="\t", quote=F)
  write.table(myDiff25p.hypo, file=paste("diff.hypo.",context,".txt",sep=""), sep="\t", quote=F)  
  write.table(myDiff25p, file=paste("diff.all.",context,".txt",sep=""), sep="\t", quote=F)
}
```

Tiling and other
```{r events, warning=FALSE, cache=TRUE}
tiles <- tileMethylCounts(myobj, win.size = 1000, step.size = 1000)
if(context==""){
  write.table(tiles, file="tiled.txt", sep="\t", quote=F)
}else{
  write.table(tiles, file=paste("tiled.",context,".txt",sep=""), sep="\t", quote=F)
}

diffMethPerChr(myDiff, plot = FALSE, qvalue.cutoff = qvalue, meth.cutoff = diff)
diffMethPerChr(myDiff, plot = TRUE, qvalue.cutoff = qvalue, meth.cutoff = diff)
if(context==""){
  write.table(myDiff, file="diff.txt", sep="\t", quote=F)
}else{
  write.table(myDiff, file=paste("diff.",context,".txt",sep=""), sep="\t", quote=F)
}

```

Save the workspace
```{r save}
if(context!=""){
  save.image(file=paste("Control_vs_LR5007_ApeKI_",context,"_", chromosome, ".RData", sep=""))
}else{
  save.image(file=paste("Control_vs_LR5007_ApeKI_", chromosome, ".RData", sep=""))
}
```
